---
title: "Eigenvalue Draft"
author: "Wells"
date: "February 15, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

 We create a matrix of $n^2$ standard Gaussian variables
```{r}
rmatrix<-function(n){
  vec<-rnorm(n^2,0,1)
  mat<-matrix(vec,n)
  return(mat)
}
rmatrix(4)
```
We compute the eigenvalues of the matrix and create a complex plot 
```{r}
randomeigen<-function(n){
  M<-rmatrix(n)
  ev<-eigen(M,only.values=TRUE)
  (values<-ev$values)
  return(values)
}
plot(randomeigen(400),asp=1) #use the asp=1 parameter to ensure a square plot
```
For comparison purposes, the graph of the complex circle of radius 1, $|z|=1$, is given by
```{r}
x<-seq(-pi,pi,0.1)
z<- complex(real=0,imag=x)
plot(exp(z),asp=1,type="l")
```
```{r}
plot(randomeigen(400),asp=1)
lines(20*exp(z))
```

How does this compare to $n$ points chosen uniformly on the disk of radius $\sqrt{n}$?

```{r}
unifdisk<-function(n){
  s<-sqrt(n)
  rho <- sqrt(runif(n))
  theta <- runif(n, 0, 2*pi)
  x <- rho *s* cos(theta)
  y <- rho *s* sin(theta)
  plot(x, y,asp=1)
}
unifdisk(400)
t<-seq(-2*pi,2*pi,0.1)
lines(20*cos(t),20*sin(t))
```

Note that eigenvalues for a matrix come in conjugate pairs, so we aren't quite making a fair comparison. Instead, we should uniformly sample $n/2$ points on the upper semicircle, and then reflect across the real axis.
```{r}
unifsemi<-function(N){
  n<-N/2
  s<-sqrt(N)
  rho <- sqrt(runif(n))
  theta <- runif(n, 0, pi)
  x1 <- rho *s* cos(theta)
  x2<- rho *s* cos(theta)
  x<- c(x1,x2)
  y1 <- rho *s* sin(theta)
  y2<- -rho *s* sin(theta)
  y<- c(y1,y2)
  plot(x, y,asp=1)
}
unifsemi(400)
t<-seq(-2*pi,2*pi,0.1)
lines(20*cos(t),20*sin(t))
```

[Discuss density function for eigenvalues?]

If we restrict our attention to those matrices with real eigenvalues, we can obtain a cleaner formula for the joint density function of those eigenvalues. Recall from the Spectral Theorem (or trust me, if you haven't seen this before) that a complex matrix $A$ has real eigenvalues if it is self-adjoint, in the sense that $A = \overline{A}^T$. In the special case when $A$ actually has real entries, then self-adjoint matrices are just symmetric matrices.

There are a few ways to obtain self-adjoint random matrices, but the simplest to execute in R is to generate an ordinary random matrix, and then average across that matrix and its conjugate transpose:

$$A =\frac{M + \overline{M}^T }{\sqrt{2}}$$

We generate a self-adjoint real matrix:
```{r}
rmatrixReal<-function(n){
  M<-rmatrix(n)
  A<-(M+t(M))/(sqrt(2)) #Real matrix with variance 1
  return(A)
}
A<-rmatrixReal(4)
A
isSymmetric.matrix(A)
```

And we generate a self-adjoint complex matrix:

```{r}
rmatrixComplex<-function(n){
  RP<-rmatrix(n) #Real Part
  IP<-rmatrix(n)*1i #Imaginary Part
  C<-(RP + IP)/sqrt(2) #Complex Matrix with variance 1
  A<-(C+t(Conj(C)))/sqrt(2)
  return(A)
}
A<-rmatrixComplex(4)
A
isSymmetric.matrix(A)

```


As before, we calculate and plot the eigenvalues of A:
```{r}
randomeigenReal<-function(n){
  M<-rmatrixReal(n) 
  ev<-eigen(M,only.values=TRUE,symmetric=TRUE)
  (values<-ev$values)
  return(values)
}
#randomeigenReal(4)
x<-randomeigenReal(64)
z<-complex(real=x, imag=0)
plot(z,asp=1)
```
```{r}
randomeigenComplex<-function(n){
  M<-rmatrixComplex(n) 
  ev<-eigen(M,only.values=TRUE,symmetric=TRUE)
  (values<-ev$values)
  return(values)
}
#randomeigenReal(4)
x<-randomeigenComplex(64)
z<-complex(real=x, imag=0)
plot(z,asp=1)
```
Finally, we can plot the location of $n$ variables choosen uniformly on the interval $[-2\sqrt{n}, 2\sqrt{n}]$
```{r}
N<-64 #Number of variables
x<-runif(N,- 2*sqrt(N), 2*sqrt(N))
z<-complex(real=x,imag=0)
plot(z,asp=1)
```

